---
title: "Predator Prey Stability"
author: "Victoria Cutler & Mallory  Giesie"
date: "2023-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sensitivity)
library(tidyverse)
library(deSolve)
library(lhs)
library(purrr)
library(ggpubr)
```

**Stability Assignment:**

Consider how you might add hunting of prey to the predator prey model that we've been using in class

You can assume the following rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000

**Exploring Hunting Levels and Minimum Prey Populations to Permit Hunting**

```{r}
source("R/lotvmodK.R")
source("R/stability_measure.R")

currpop = c(prey=70, pred=50) #Defining our starting population 

# looking at hunting levels and prey populations
hunt_level <- seq(from = 0, to = 1, by = 0.01)
dpredprey <- map_df(hunt_level, ~lotvmodK(pop = currpop, t = NULL, pars = list(rprey = 0.95, alpha = 0.01, eff = 0.6, pmort = 0.4, K = 2000, min_prey_pop = 70, hunt_level = .x)))

a = ggplot(data = dpredprey, aes(hunt_level, prey_growth))+geom_point()+geom_hline(yintercept = 0, col="red")+
  labs(y="Prey Growth Rate", x="Hunting Rate")
a 

print("If we assume that predator and prey populations are at a value of 70 & 50 (and rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000), in order to maintain a postive growth rate for prey, with a minimum prey population to hunt of 70, the hunting rate should exceed more than 40%")
```

**We defined our stability metric under this scenario: suppose the predator is an endangered species, which there is a goal that in 100 years, the population should be twice the size of the current population. Under this scenario, both our goal and our measure of stability will be growth**

*Therefor, our stability metric is that the population of the predator must be twice the size of the starting condition*

```{r}
# set parameter list
pars = data.frame(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000, hunt_level = 0.10, min_prey_pop = 80)
# times when you want to evaluate
days = seq(from=1,to=100)
res = ode(func=lotvmodK, y=currpop, times=days, parms=pars)
res1 = as.data.frame(res) %>% gather(key="animal", value="pop",-time)
p1 = ggplot(res1, aes(time, pop, col=animal))+geom_line()
p1

stability <- stability_measure(res1)
print(paste("The system is stable: ", stability))


# set parameter list
pars = data.frame(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000, hunt_level = 0.20, min_prey_pop = 80)
# times when you want to evaluate
days = seq(from=1,to=100)
res = ode(func=lotvmodK, y=currpop, times=days, parms=pars)
res2 = as.data.frame(res) %>% gather(key="animal", value="pop",-time)
p2 = ggplot(res2, aes(time, pop, col=animal))+geom_line()
p2

stability <- stability_measure(res2)
print(paste("The system is stable: ", stability))


# set parameter list
pars = data.frame(rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000, hunt_level = 0.05, min_prey_pop = 110)
# times when you want to evaluate
days = seq(from=1,to=100)
res = ode(func=lotvmodK, y=currpop, times=days, parms=pars)
res3 = as.data.frame(res) %>% gather(key="animal", value="pop",-time)
p2 = ggplot(res3, aes(time, pop, col=animal))+geom_line()
p2

stability <- stability_measure(res3)
print(paste("The system is stable: ", stability))
```

**Recommended Hunting Strategy**

*To bring our system back to stability (or double our predator population) we recommend that the hunting level doesn't exceed 5%, bearing in mind that you should let the prey population grow to a population of 110 before you let hunting begin. This is necessary because we need there to be abundant prey for the predators to eat to then grow their population*

